---
title: "Continuity Example"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Continuity Example}
 %\VignetteIndexEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE, warning=FALSE);

library(stringi, include.only = "%s+%");
library(magrittr);
library(purrr);
library(event.vectors);

source("list2html.R");
htmltools::tags$style(readLines("markdown.css") |> paste(collapse = "\n")) |> htmltools::tagList();

make.test_data <- function(j = 5, n = 5, m = 5, o = 1:10, dest = globalenv(), .debug = FALSE){
#' Make Test Data for Validation
#'
#' \code{make.test_data} creates several \code{\link[data.table]{data.table}} objects to be used to validate the package functionality
#' @param j (integer | 5L) The number of unique values for 'k' to generate
#' @param n (integer | 5L) The number of sources to create (maximum of 10)
#' @param m (integer | 5L) The maximum number of columns to generate for each created object (maximum of 15): each column represents an attribute not related to time
#' @param o (integer[] | 1:10) A vector of integers sampled to create end dates from a randomly-generated start date
#' @param dest (environment) The destination environment object
#' @param .debug (logical | \code{FALSE}) When \code{TRUE}, additional debugging items are printed
#'
#' @return One to \code{n} \code{\link[data.table]{data.table}} objects prefixed as 'test_data'.
#'
	j = max(c(3, abs(as.integer(j))));
	n = max(c(3, abs(as.integer(n)))); ifelse(n > 10, 10, n);
	m = max(c(3, abs(as.integer(m)))); ifelse(m > 15, 15, m);
	o = { ifelse(
					rlang::has_length(o, 1L)
					, yes = 1:10
					, no = ifelse(
							any(o <= 0)
						 	, yes = abs(o)
						 	, no = ifelse(
						 			rlang::has_length(o, 2L)
						 			, yes = `:`(o[1], o[2]) |> sort()
						 			, no =  o
						 			))
				)
		}

	sequence(n) %>%
	purrr::set_names(paste0("test_data.", stringi::stri_pad_left(., width = 2, pad = "0")))|>
	purrr::map(~{
		set.seed(90210);
		.src = LETTERS[[.x]];

		.out = purrr::map_dfr(c(1:j), ~list(join_key = rep.int(.x, sample(10:100, 1, TRUE)), src = .src)) %>% data.table::as.data.table();

		.init_date = c(as.Date(sprintf(
				"%s-%s-%s"
				, rep.int(data.table::year(Sys.Date()), nrow(.out))
				, sample(stringi::stri_pad_left(1:12, width = 2, pad = "0"), nrow(.out), TRUE)
				, sample(stringi::stri_pad_left(1:28, width = 2, pad = "0"), nrow(.out), TRUE)
				)));

		.out[
			, c("date.start", "date.end") := list(.init_date, .init_date + rpois(n = length(join_key), lambda = sample(o, length(join_key), TRUE)))
			][
			, paste0("X_", stringi::stri_pad_left(sample(30, m), width = 2, pad = "0")) := purrr::map(1:m, ~sample(runif(1E6), .N, TRUE))
			][
			runif(length(join_key)) > 0.65
			] %>%
			data.table::setkey(join_key, src, date.start, date.end) %>%
			data.table::setcolorder(c("join_key", "date.start", "date.end", "src"))
		}) %>%
	list2env(envir = dest);
}
make.test_data(j = 50, n = 1, m = 5, o = c(5, 20), dest = .GlobalEnv, .debug = !TRUE);

```

## Synthesized Data 

```{r SYNTH_DATA}
test_data.01[, join_key:src] |>
	DT::datatable(
		options = list(dom = "rtip")
		, rownames = FALSE
		, extensions = c("Responsive", "Scroller")
		) |>
	DT::formatStyle(columns =1:4, textAlign = "center")
```


## Continuity Results
Using a `timeout` of 10 (days) and imputed groups defined by columns `join_key` and `src`, the data was consolidated into a series of larger epochs marked by _start_ and _end_ dates. The data will be processed under.

```{r CONTINUITY_RUN, warning=FALSE}
data.table::copy(test_data.01) |>
	continuity(
		map_fields = c(join_key, src)
		, time_fields = c(date.start, date.end)
		, boundary_name = episode
		, timeout = 10
		, archipelago = TRUE
		, show.all = !TRUE
		) |>
	DT::datatable(
		options = list(dom = "rtip")
		, rownames = FALSE
		, extensions = c("Responsive", "Scroller")
		) |>
	DT::formatStyle(columns = 1:5, textAlign = "center")
```

## A Final Note

There is an additional consideration that is not addressed by the `continuity` function: Is the timeout choice of `10` the best choice given the data?  Answering that requires comparing imputed "islands" to some independent, adjudicating set of data or domain expert that can provide an assessment of reasonableness.  An alternative method would be to set the timeout by means of statistical analysis.  See [Break Signal](Break-Signal.html) for an example.